### 2025-06-09 simulation.qmd
20250609 - bootstrapping

# Simulation: the Monte Carlo Method {#sec-simulation}

@@

```{r}
library("tidyverse")
```

@@

```{r}
load(file = "./data/players_projectedPoints_seasonal.RData")
```

@@

@@

df = ffanalytics:::source_points(l, scoring_rules = ffanalytics::scoring)
```

@@

```{r}
calculate_fantasyPoints <- function(df, scoring_rules){
  #fantasyPoints <-

>>>>

# Simulation: Bootstrapping and the Monte Carlo Method {#sec-simulation}

@@

```{r}
library("ffanalytics")
library("data.table")
library("future")
library("future.apply")
library("tidyverse")
```

@@

```{r}
#| eval: false
#| include: false

load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/players_projectedPoints_seasonal.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_playerIDs.RData", fsep = ""))
```

@@

```{r}
all_proj <- bind_rows(players_projectedPoints_seasonal)
```

```{r}
vars_by_pos <- list(
  QB = c(
    "games",
    "pass_att", "pass_comp", "pass_yds", "pass_tds", "pass_int",
    "rush_att", "rush_yds", "rush_tds",
    "fumbles_lost"#,
    #"sacks",
    #"pass_09_tds", "pass_1019_tds", "pass_2029_tds", "pass_3039_tds", "pass_4049_tds", "pass_50_tds",  
    #"pass_250_yds", "pass_300_yds", "pass_350_yds", "pass_40_yds", "pass_400_yds",
    #"rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds"
    ),
  RB = c(
    "games",
    "rush_att", "rush_yds", "rush_tds",
    "rec_tgt", "rec", "rec_yds", "rec_tds",
    "fumbles_lost",
    "return_yds", "return_tds"#,
    #"rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
    #"rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
    #"rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds",
  ),
  WR = c(
    "games",
    "pass_att", "pass_comp", "pass_yds", "pass_tds", "pass_int",
    "rush_att", "rush_yds", "rush_tds",
    "rec_tgt", "rec", "rec_yds", "rec_tds",
    "fumbles_lost",
    "return_yds", "return_tds"#,
    #"rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
    #"rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
    #"rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds",
  ),
  TE = c(
    "games",
    "pass_att", "pass_comp", "pass_yds", "pass_tds", "pass_int",
    "rush_att", "rush_yds", "rush_tds",
    "rec_tgt", "rec", "rec_yds", "rec_tds",
    "fumbles_lost",
    "return_yds", "return_tds"#,
    #"rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
    #"rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
    #"rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds",
  )
)
```

```{r}
bootstrapSimulation <- function(projectedStats, vars_by_pos, n_iter = 10000, seed = NULL) {
  dt <- data.table::as.data.table(all_proj)
  all_ids <- unique(dt$id)

  if (!is.null(seed)) set.seed(seed)

  future::plan(multisession)

  results <- future.apply::future_lapply(
    all_ids,
    function(player_id) {
      player_data <- dt[id == player_id]
      player_pos <- unique(player_data$pos)

      if (length(player_pos) != 1 || !player_pos %in% names(vars_by_pos)) return(NULL)

      stat_vars <- vars_by_pos[[player_pos]]
      out <- data.table::data.table(iteration = 1:n_iter, id = player_id, pos = player_pos)

      for (var in stat_vars) {
        if (var %in% names(player_data)) {
          non_na_values <- player_data[[var]][!is.na(player_data[[var]])]

          if (length(non_na_values) > 0) {
            out[[var]] <- sample(non_na_values, n_iter, replace = TRUE)
          } else {
            out[[var]] <- NA_real_
          }
        } else {
          out[[var]] <- NA_real_
        }
      }

      return(out)
    },
    future.seed = TRUE
  )

  data.table::rbindlist(results, use.names = TRUE, fill = TRUE)
}

bootstappedStats <- bootstrapSimulation(
  projectedStats = all_proj,
  vars_by_pos = vars_by_pos,
  n_iter = 10000,
  seed = 52242)

bootstappedStats$data_src <- bootstappedStats$iteration

bootstappedStatsByPosition <- split(
  bootstappedStats,
  by = "pos",
  keep.by = TRUE)

bootstappedStatsByPosition <- lapply(
  bootstappedStatsByPosition,
  as_tibble)

attr(bootstappedStatsByPosition, "season") <- 2024
attr(bootstappedStatsByPosition, "week") <- 0

bootstrappedFantasyPoints <- ffanalytics:::source_points(
  data_result = bootstappedStatsByPosition,
  scoring_rules = ffanalytics::scoring)

bootstrappedFantasyPoints$iteration <- bootstrappedFantasyPoints$data_src
bootstrappedFantasyPoints$data_src <- NULL

bootstrappedFantasyPoints <- bootstrappedFantasyPoints %>% 
  left_join(
    nfl_playerIDs[,c("mfl_id","name","merge_name","team")],
    by = c("id" = "mfl_id")
  )
```

```{r}
#| eval: false
#| include: false

@@

df <- ffanalytics:::source_points(data_result = l, scoring_rules = ffanalytics::scoring)
```

```{r}
bootstrappedFantasyPoints_summary <- bootstrappedFantasyPoints %>% 
  group_by(id) %>% 
  summarise(
    mean = mean(raw_points, na.rm = TRUE),
    SD = sd(raw_points, na.rm = TRUE),
    min = min(raw_points, na.rm = TRUE),
    max = max(raw_points, na.rm = TRUE),
    range = max(raw_points, na.rm = TRUE) - min(raw_points, na.rm = TRUE),
    IQR = IQR(raw_points, na.rm = TRUE),
    MAD = mad(raw_points, na.rm = TRUE),
    CV = SD/mean,
    median = median(raw_points, na.rm = TRUE),
    pseudomedian = DescTools::HodgesLehmann(raw_points, na.rm = TRUE),
    mode = petersenlab::Mode(raw_points, multipleModes = "mean"),
    skewness = psych::skew(raw_points, na.rm = TRUE),
    kurtosis = psych::kurtosi(raw_points, na.rm = TRUE)
  )

bootstrappedFantasyPoints_summary <- bootstrappedFantasyPoints_summary %>% 
  left_join(
    nfl_playerIDs[,c("mfl_id","name","merge_name","team","position")],
    by = c("id" = "mfl_id")
  ) %>% 
  select(name, team, position, mean:kurtosis, everything())

bootstrappedFantasyPoints_summary
```

bootstrappedFantasyPoints_summary %>% filter(mean > 100 & position == "QB") %>% arrange(CV) %>% head
bootstrappedFantasyPoints_summary %>% filter(mean > 100 & position == "QB") %>% arrange(-CV) %>% head
```

```{r}
#| label: fig-bootstrapping1
#| fig-cap: "Distribution of Projected Fantasy Points for Patrick Mahomes from Bootstrapping."
#| fig-alt: "Distribution of Projected Fantasy Points for Patrick Mahomes from Bootstrapping."

ggplot2::ggplot(
  data = bootstrappedFantasyPoints %>%
    filter(pos == "QB" & name == "Patrick Mahomes"),
  mapping = aes(
    x = raw_points)
) +
  geom_histogram(
    aes(y = after_stat(density)),
    color = "#000000",
    fill = "#0099F8"
  ) +
  geom_density(
    color = "#000000",
    fill = "#F85700",
    alpha = 0.6 # add transparency
  ) +
  geom_rug() +
  #coord_cartesian(
  #  xlim = c(0,400)) +
  labs(
    x = "Fantasy Points",
    y = "Density",
    title = "Distribution of Projected Fantasy Points for Patrick Mahomes"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

```{r}
#| label: fig-bootstrapping1
#| fig-cap: "Distribution of Projected Fantasy Points for Two Players from Bootstrapping. There is relatively narrow uncertainty around projected fantasy points for Dak Prescott, whereas there is relatively wide uncertainty around the projected fantasy points for Drake Maye."
#| fig-alt: "Distribution of Projected Fantasy Points for Two Players from Bootstrapping. There is relatively narrow uncertainty around projected fantasy points for Dak Prescott, whereas there is relatively wide uncertainty around the projected fantasy points for Drake Maye."

ggplot2::ggplot(
  data = bootstrappedFantasyPoints %>%
    filter(pos == "QB" & (name %in% c("Dak Prescott", "Drake Maye"))),
  mapping = aes(
    x = raw_points,
    group = name,
    #color = name,
    fill = name)
) +
  geom_histogram(
    aes(y = after_stat(density))
  ) +
  geom_density(
    alpha = 0.6 # add transparency
  ) +
  coord_cartesian(
    xlim = c(0,NA),
    expand = FALSE) +
  #geom_rug() +
  labs(
    x = "Fantasy Points",
    y = "Density",
    fill = "",
    color = "",
    title = "Distribution of Projected Fantasy Points"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

@@

```{r}
#| eval: false
#| include: false

calculate_fantasyPoints <- function(df, scoring_rules){
  #fantasyPoints <-
